\chapter{Evaluation}
To verify our work on the convergence layer as well as compare it to other and
find potential bottlenecks we evaluated out work under different test cases.
Comparing the data throughput and packetloss for raw IEEE 802.15.4 against DTN
as an additional layer. Distance measurement as well as Round Trip Time
measurement was also done.

\section{Throughput}

\section{Packetloss}

\section{Distance}
The Imote2 has no connector for an external antenna but an on-board antenna
only. To get an understanding what use cases this antenna, and therefor the
whole board, could cover we combined our measurements with a distance
measurement. To decide if a distance is still usable enough we used the
packetloss as quality indicator. All distances which showed a packetloss above
50 percent are rated as unusable. During our measurements such a high packetloss
was shown for all distances higher then 33m.


\section{DTN Round Trip Time}
The last testcase was the round trip time of a DTN bundle. We used the
\emph{dtnping} utility to send out a bundle, receive the answer and calculate
the round trip time. Bundles are send to the \emph{echo} application of the
receiving node. In our case this is directly the DTN daemon. The default payload
size of a ping bundle was to big to our limited payload size on the IEEE
802.15.4 convergence layer. We reduced the payload to 1 byte and used it for all
tests. Table \ref{dtnrtt} shows the results of the measurement in milliseconds. We
had 10 different scenarios for this test case and did five measurements for each.
The first eight are the same we used for our throughput and packetloss test cases.
We added two more scenarios to compare it against our usual test environment. The
local scenario did send bundles over the UDP convergence layer on the loopback
interface of our X86 workstation while the UDP scenario did send the bundles
through the UDP convergence layer over an 1 Gbit Ethernet link to another
machine.

\begin{table}
\begin{tabular}{l*{9}{c}r}
    & 0m & 1m & 5m & 10m & 15m & 20m & 25m & 30m & local & UDP \\
\hline
Min & 656.14 & 651.11 & 664.25 & 645.27 & 659.29 & 651.40 & 653.66 & 677.29 & 38.31 & 40.90 \\
Mean & 676.17 & 662.38 & 674.85 & 674.85 & 676.46 & 672.03 & 672.47 & 686.60 & 38.87 & 41.26 \\
Max & 693.14 & 678.49 & 684.96 & 685.11 & 695.76 & 684.22 & 687.26 & 693.64 & 39.41 & 41.56 \\
\end{tabular}
\caption{ DTN Round Trip Time measurement results}
\label{dtnrtt}
\end{table}
